\chapter*{Acknowledgements}

My pursuit of a master's degree was a challenging part-time undertaking that began in 2018 and lasted over five years. During this time, the world encountered a global crisis of capitalism and the widespread COVID-19 pandemic, which had a significant impact on everyone. I switched jobs, and at times, I contemplated abandoning this study. However, I am very lucky for the support of many individuals who made it possible for me to complete this endeavor.

First, I express my gratitude to my advisor, Prof. Denis Mauá, for providing me with invaluable guidance and motivation throughout my academic journey. Our numerous lengthy meetings were filled with insights, recommendations of research papers, ideas for experiments and constructive feedback on my work. I am thankful for his understanding of my time constraints and his dedication to throughly reviewing my work.

Furthermore, I extend my thanks to Prof. Denis' students --- Heitor Ribeiro, Julissa Llerena, and Renato Geh --- for generously sharing their texts and codes with me. Their contributions were instrumental in enhancing my comprehension of probabilistic models and facilitating the experiments conducted in this study.

I would like to express my appreciation to Prof. Alessandro Antonucci and Prof. Fabio Cozman for accepting the invitation to compose this work's examining commitee, and to Prof. Glauber de Bona and Prof. Marcelo Queiroz for their feedback during my qualifying exam.

A special thank you to my friend, Prof. David Kohan, for providing me with constant academic encouragement. Our coffee meetings over random mathematical puzzles and the annoying question ``How is the master's going?'' were crucial for me to complete this work, as well as an invitation to write a paper on a different topic a couple of years ago.

I am very grateful for the exceptional education I received from the Institute of Mathematics and Statistics at the University of São Paulo (IME-USP), a public institution where I have been a student for almost 15 years.

Without music and friends nothing would make sense. Big thanks to two great musicians of our time, Kiko Horta and Toninho Ferragutti, for the accordion lessons, which were a key fuel to living through the past years, as were the musical meetings of Forró da Varanda and many others.

Finally, I express my heartfelt gratitude to Juliana for the everyday partnership, and to my parents, Amarildo and Márcia, and my brothers, Bruno and Lucas, for their lifelong love and support.
